---
title: "Optimizing an Investment Portfolio using Nonlinear Programming"
author: "Kelly Ng"
date: "2024-02-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r}
library(nloptr)
library(quantmod)
library(tseries)
```

### Set the date range for historical price data
```{r}
start_date <- "2013-03-31"
end_date <- "2023-03-31"
```

### Fetch historical price data
```{r}
aapl_prices <- getSymbols("AAPL", src = "yahoo", from = start_date, to = end_date, auto.assign = FALSE)
msft_prices <- getSymbols("MSFT", src = "yahoo", from = start_date, to = end_date, auto.assign = FALSE)
amzn_prices <- getSymbols("AMZN", src = "yahoo", from = start_date, to = end_date, auto.assign = FALSE)
nvda_prices <- getSymbols("NVDA", src = "yahoo", from = start_date, to = end_date, auto.assign = FALSE)
googl_prices <- getSymbols("GOOGL", src = "yahoo", from = start_date, to = end_date, auto.assign = FALSE)
```

### Calculate daily returns
```{r}
aapl_returns <- dailyReturn(Ad(aapl_prices))
msft_returns <- dailyReturn(Ad(msft_prices))
amzn_returns <- dailyReturn(Ad(amzn_prices))
nvda_returns <- dailyReturn(Ad(nvda_prices))
googl_returns <- dailyReturn(Ad(googl_prices))
```

### Merge daily returns
```{r}
all_returns <- merge(aapl_returns, msft_returns, amzn_returns, nvda_returns, googl_returns)
```

### Compute the covariance matrix
```{r}
cov_matrix <- cov(all_returns)
cov_matrix
```

### Mean returns 
```{r}
mean_returns <- colMeans(all_returns)
confidence_level <- 0.95
alpha <- 0.5
n_assets <- length(mean_returns)
```

### Objective function
```{r}
objective_function <- function(weights, alpha, mean_returns, cov_matrix, confidence_level) {
  expected_return <- sum(mean_returns * weights)
  portfolio_variance <- as.numeric(t(weights) %*% cov_matrix %*% weights)
  portfolio_volatility <- sqrt(portfolio_variance)
  z_score <- qnorm(1 - confidence_level)
  cvar <- portfolio_volatility * z_score / (1 - confidence_level)
  objective_value <- alpha * expected_return - (1 - alpha) * cvar
  return(-objective_value)
}
```

### Equality constraint: sum of the weights should be equal to 1 (budget constraint)
```{r}
equality_constraint <- function(weights) {
  return(sum(weights) - 1)
}
```

### Gradient function
```{r}
gradient_function <- function(weights, alpha, mean_returns, cov_matrix, confidence_level) {
  # Gradient of return
  grad_expected_return <- mean_returns
  
  # Gradient of CVaR
  portfolio_variance <- as.numeric(t(weights) %*% cov_matrix %*% weights)
  portfolio_volatility <- sqrt(portfolio_variance)
  z_score <- qnorm(1 - confidence_level)
  grad_cvar <- cov_matrix %*% weights * z_score / (portfolio_volatility)
  
  # Gradient of the objective function
  grad_objective_value <- alpha * grad_expected_return - (1 - alpha) * grad_cvar
  return(grad_objective_value)
}
```

### Wrapper functions
```{r}
objective_function_wrapper <- function(alpha, mean_returns, cov_matrix, confidence_level) {
  function(weights) {
    objective_function(weights, alpha, mean_returns, cov_matrix, confidence_level)
  }
}

equality_constraint_wrapper <- function() {
  function(weights) {
    equality_constraint(weights)
  }
}

gradient_function_wrapper <- function(alpha, mean_returns, cov_matrix, confidence_level) {
  function(weights) {
    gradient_function(weights, alpha, mean_returns, cov_matrix, confidence_level)
  }
}
```

### Generate random initial weights
```{r}
set.seed(123)  # Set seed for reproducibility
initial_weights <- runif(n_assets)
initial_weights <- initial_weights / sum(initial_weights)  # Normalize so the sum of weights is 1
```

### Optimize the portfolio weights using the nloptr function
```{r}
gradient_equality_constraint_wrapper <- function(x) {
  # Assuming x is your vector of decision variables (weights in this case)
  # And the constraint is sum(weights) = 1, the gradient is a vector of ones
  return(rep(1, length(x)))
}

result <- nloptr(x0 = initial_weights,
                 eval_f = objective_function_wrapper(alpha, mean_returns, cov_matrix, confidence_level),
                 eval_grad_f = gradient_function_wrapper(alpha, mean_returns, cov_matrix, confidence_level),
                 lb = rep(0, n_assets),  # All weights should be non-negative (no short-selling)
                 ub = rep(1, n_assets),  # All weights should be less than or equal to 1
                 eval_g_eq = equality_constraint_wrapper(),  # Ensure the sum of weights is equal to 1
                 eval_jac_g_eq = gradient_equality_constraint_wrapper,  # Gradient for equality constraint
                 opts = list(algorithm = "NLOPT_LD_SLSQP", maxeval = 1000, xtol_rel = 1e-6))
result
```

